Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-07-15 Master Thesis/synopsis.tex
Add more bibliography items

\begin{definition}\label{2.1}
Let $K$ be a field of finite dimension. We denote the set of equivalence classes of Krull valuations of $K$ by $\operatorname{Spv}(K)$. We often do not distinguish between a valuation and its equivalence class.

@@

\begin{definition}
\[\{v\in\operatorname{Spv}(K)\mid v(x)\geq 0\text{ for all }x\in S\} \]
We can endow $\operatorname{Spv}(K)$ with a natural topology by declaring the following sets as a basis of open sets:
where $S$ ranges over all finite subsets of $K$. 
\end{definition}

\begin{lemma}\label{spv-topology}
$\operatorname{Spv}(K)$ with the topology defined above is a quasi-compact $T_0$-space. In particular the set of closed points $|\operatorname{Spv}(K)|$ of $\operatorname{Spv}(K)$ is a quasi-compact $T_1$-space, endowed with the induced topology. Let $v,w\in\operatorname{Spv}(K)$. We have:
\[\mathcal{O}_w\subset\mathcal{O}_v \iff w\in\overline{\{v\}} \]
Thus $v$ is a closed point if and only if it has no non-trivial specializations.
\end{lemma}

\begin{proof}
See \cite{ZS60} VI §17 Theorem 38, §17 Theorem 40.
\end{proof}

@@

\begin{proposition}[\textit{Kummer Theory}]\label{prop:kummer-theory}
Let $F$ be a field and $n\in\mathbb{N}$ invertible in $F$. Let $\mu_n\subset F^\text{s}$ be the group of all $n$-th roots of unity. Consider the following short exact sequence:
\[ 1\longrightarrow \mu_n\longrightarrow (F^\text{s})^\times \stackrel{n}{\longrightarrow} (F^\text{s})^\times \longrightarrow 1 \]
This gives rise to the following long exact cohomology sequence:
\[\begin{tikzcd}
\cdots\longrightarrow F^\times \ar[r, "n"]& F^\times \ar[r]& \mathrm{H}^1(F, \mu_n)\ar[r] & \mathrm{H}^1(F, (F^\text{s})^\times)\ar[dlll]\\
\mathrm{H}^1(F, (F^\text{s})^\times)\ar[r] & \mathrm{H}^2(F, \mu_n)\ar[r] & \operatorname{Br}(F)\ar[r, "n"] & \operatorname{Br}(F) \longrightarrow \cdots
\end{tikzcd}\]
Since $\mathrm{H}^1(F, (F^\text{s})^\times)=0$ by Hilbert's Theorem 90, we get isomorphisms:
\begin{align*}
\mathrm{H}^1(F,\mu_n) \cong F^\times/n, \qquad \mathrm{H}^2(F, \mu_n) \cong {}_n\operatorname{Br}(F)
\end{align*}
\end{proposition}

@@

The second assertion follows from the first and from [K2], §5, Theorem 1.
\end{proof}

@@

\begin{proof}
(i) and (ii) follow from [HK], Lemma 2.4.3 and (iii) is a consequence of (ii) and the fact that henselizations commute with filtrated direct limits, see \cite{EGA4} 18.6.14.
\end{proof}

@@

In this section we shall show that $\sigma: G_K\to G_{K'}$ induces group isomorphisms $\varphi_n: K'^\times/n\to K^\times/n$.

Let $k$ and $k'$ be the algebraic closures of $\mathbb{Q}$ in $K$ and $K'$ respectively, i.e. the constant fields of $K$ and $K'$, and let $\overline{k}$ and $\overline{k'}$ be the algebraic closure of $k$ and $k'$ in $\overline{K}$ and $\overline{K'}$ respectively.

@@

By Lemma \ref{3.2} $\sigma$ induces a bijection $\sigma: X^{\phantom{'}}_0\to X'_0,\ P\mapsto P'$ and for any $P\in X_0$ an isomorphism $\varphi_P: k(P')\to k(P)$.

@@

\begin{definition}\label{4.4}
For a finite subset $S\subset X_0$ set:
\[ K^S=\{x\in K^\times \mid v_P(x)=0\text{ for all }P\in X_0 -  S \} = \bigcap_{P\in X_0 -  S} U_P \]
$K'^{S'}$ is defined similarly. For finite subsets $S$ of $X_0$, the corresponding subsets of $X_0'$ will be denoted by $S'$, i.e. $S'=\{\sigma P\mid P\in S \} $.
\end{definition}

\begin{theorem}[\textit{Weak Approximation Theorem}]\label{weak-approximation}
Let $P_1,\ldots,P_n\in X_0$ be pairwise distinct closed points, $x_1,\ldots,x_n\in K$ and $a_1,\ldots,a_n\in\mathbb{Z}$. Then there is some $x\in K$, such that for all $i$ we have:
\[ v_{P_i}(x-x_i) = a_i \]
\end{theorem}

>>>>

\paragraph{} Let $K$ be a field of finite dimension.

\begin{definition}\label{2.1}
We denote the set of equivalence classes of Krull valuations of $K$ by $\operatorname{Spv}(K)$. We often do not distinguish between a valuation and its equivalence class.

@@

\begin{definition}
We can endow $\operatorname{Spv}(K)$ with a natural topology by declaring the following sets as a basis of open sets:
\[U_S =\{v\in\operatorname{Spv}(K)\mid v(x)\geq 0\text{ for all }x\in S\} \]
where $S$ ranges over all finite subsets of $K$. 
\end{definition}

\begin{lemma}\label{spv-topology}
Let $v,w\in\operatorname{Spv}(K)$. We have:
\[\mathcal{O}_w\subset\mathcal{O}_v \iff w\in\overline{\{v\}} \]
Thus $v$ is a closed point if and only if it has no non-trivial specializations.

$\operatorname{Spv}(K)$ is a quasi-compact Jacobson space, i.e. every closed subset $A\subset\operatorname{Spv}(K)$ is the closure of its closed points. In particular, the set of closed points $|\operatorname{Spv}(K)|$ of $\operatorname{Spv}(K)$ is a quasi-compact $T_1$-space, endowed with the induced topology. 
\end{lemma}

\begin{proof}
The first assertion follows from \cite{ZS60} VI, §17 Theorem 38. The quasi-compactness of $\operatorname{Spv}(K)$ follows from \cite{ZS60} VI, §17 Theorem 40.

Let $A\subset \operatorname{Spv}(K)$ be a closed set, $S\subset K$ finite, and assume $U_S\cap A\neq\varnothing$. Let $v\in U_S\cap A$ be a point which is not closed.
\end{proof}

@@

\begin{proposition}\label{prop:kummer-theory}
Let $F$ be a field and $n\in\mathbb{N}$ invertible in $F$. Let $\mu_n\subset F^\text{s}$ be the group of all $n$-th roots of unity. Consider the following short exact sequence:
\[ 1\longrightarrow \mu_n\longrightarrow (F^\text{s})^\times \stackrel{n}{\longrightarrow} (F^\text{s})^\times \longrightarrow 1 \]
This gives rise to the following long exact cohomology sequence:
\[\begin{tikzcd}
\cdots\longrightarrow F^\times \ar[r, "n"]& F^\times \ar[r]& \mathrm{H}^1(F, \mu_n)\ar[r] & \mathrm{H}^1(F, (F^\text{s})^\times)\ar[dlll]\\
\mathrm{H}^1(F, (F^\text{s})^\times)\ar[r] & \mathrm{H}^2(F, \mu_n)\ar[r] & \operatorname{Br}(F)\ar[r, "n"] & \operatorname{Br}(F) \longrightarrow \cdots
\end{tikzcd}\]
Since $\mathrm{H}^1(F, (F^\text{s})^\times)=0$ by Hilbert's Theorem 90, we get isomorphisms:
\begin{align*}
\mathrm{H}^1(F,\mu_n) \cong F^\times/n, \qquad \mathrm{H}^2(F, \mu_n) \cong {}_n\operatorname{Br}(F)
\end{align*}
\end{proposition}

@@

The second assertion follows from the first and from \cite{Ka79}, §5, Theorem 1.
\end{proof}

@@

\begin{proof}
(i) and (ii) follow from \cite{HK94}, Lemma 2.4.3, and (iii) is a consequence of (ii) and the fact that henselizations commute with filtrated direct limits, e.g. see \cite{EGA4} 18.6.14.
\end{proof}

@@

In this section we shall show that $\sigma: G_K\to G_{K'}$ induces group isomorphisms $\varphi_n: K'^\times/n\to K^\times/n$. Due to Proposition \ref{prop:kummer-theory}, we have isomorphisms:
\[ \mathrm{H}^1(K',\mu_n)\cong K'^\times/n,\qquad \mathrm{H}^1(K,\mu_n)\cong K^\times/n \]
Our goal is to establish an isomorphism between the corresponding cohomology groups.

\begin{definition}
Let $k$ and $k'$ be the algebraic closures of $\mathbb{Q}$ in $K$ and $K'$ respectively, i.e. the constant fields of $K$ and $K'$, and let $\overline{k}$ and $\overline{k'}$ be the algebraic closure of $k$ and $k'$ in $\overline{K}$ and $\overline{K'}$ respectively.

@@

By Lemma \ref{3.2} $\sigma$ induces a bijection $\sigma: X^{\phantom{'}}_0\to X'_0,\ P\mapsto P'$ and for any $P\in X_0$ an isomorphism $\varphi_P: k(P')\to k(P)$.
\end{definition}

@@

\begin{definition}\label{4.4}
For a finite subset $S\subset X_0$ set:
\begin{align*}
K^S &= \bigcap_{P\in X_0 -  S} U_P
\end{align*}
$K'^{S'}$ is defined similarly. For finite subsets $S$ of $X_0$, the corresponding subsets of $X_0'$ will be denoted by $S'$, i.e. $S'=\{\sigma P\mid P\in S \} $.
\end{definition}

\begin{theorem}[\textit{Weak Approximation Theorem}]\label{weak-approximation}
Let $v_1,\ldots,v_n\in X_0$ be inequivalent discrete valuations, $x_1,\ldots,x_n\in K$ and $a_1,\ldots,a_n\in\mathbb{Z}$. Then there is some $x\in K$, such that for all $i$ we have:
\[ v_i(x-x_i) = a_i \]
\end{theorem}
